### 2016-02-08 prepReport/intro.tex
Structured final document

\chapter{Introduction}\label{chap:intro}

\section{Background}

\section{Objectives}

\section{Motivation}