### 2016-02-08 prepReport/workPlan.tex

\chapter{Work Plan}\label{chap:workPlan}

\section{Main Tasks}

\section{Approach}

\section{Tentative Scheduling}

\section{Software and Hardware Resources to be used}