### 2016-02-08 prepReport/stateArt.tex
Finished Teoretical Part. Started State-of-the-art. Cleaned up uncessary inital chaps

\chapter{State of the Art}\label{chap:stateArt}

\section{LSTM Applications (non-FPGA related)}

\section{Hardware Implementations of LSTM}

\section{Training Algorithms and Implementation}

\section{Final Overview}