### 2016-02-11 prepReport/intro.tex
Report Completed. Needs further polishing

\chapter{Introduction}\label{chap:intro}
Before start reading this report, let me provide you with a quick overview of its mains parts.

In Chapter~\ref{chap:intro}, I will present a conceptual background to the problem at hand, exposing the background that surrounds it~\ref{sec:intro_back}, the motivation for the desired solution~\ref{sec:intro_mot}, the objectives of it~\ref{sec:intro_obj} and, lastly, the people that will help me achieve it~\ref{sec:intro_people}.

In Chapter~\ref{chap:theorBack}, the theoreical foundations will be layed out. Section~\ref{sec:tb_basicCon} presents the basic concepts of Machine Learning for the unsuspicious reader that is not versed in that area of knowledge, and in the following sections, the theoretical details of ANNs, RNNs and LSTMs are presented, in consecutive order. Section~\ref{sec:training_lstm} provides a quick explanation of the training algorithm that I will use in the final solution, that will be outlined in Section~\ref{sec:propSol}.

In Chapter~\ref{chap:stateArt}, the state of the art of LSTMs, their applications~\ref{sec:sa_apps}, their hardware implementations~\ref{sec:sa_hardware} and the current work regarding the chosen training algorithm~\ref{sec:sa_training} are presented.

In Chapter~\ref{chap:workPlan}, the work plan for the Dissertation work is exhaustively detailed in \emph{tasks} in Section~\ref{sec:wp_tasks_td}, and their temporal arrangement is pictured in a Gantt Chart in Section~\ref{sec:wp_tasks_gantt}. Furthermore, the description of the Hardware and Software resources that I will be utilizing during the course of my work is done in Section~\ref{sec:wp_sh}

Finally, in Chapter~\ref{chap:earlyWork} I will present some of the work that I have already done to prepare this thesis.

\section{Background}\label{sec:intro_back}
The Artificial Neural Networks are one of the most popular learning structures in the field of Machine Learning. As the name, itself, suggests, their operation is inspired by the operation of the building blocks of our brain, the \textbf{neurons}. In spite of its high performance, one of their shortcomings is the fact that they cannot retain temporal dependences among incoming data samples, thus not beign suitable to process time-serie data, such as audio, video or other kinds of time-varying data streams, where current inputs have a high temporal dependence with previous and future inputs.

To address this issue, several algorithms have been used, such as Hidden Markov Models (HMM) or Recurrent Neural Networks (RNN), but both of these methods fail to recall temporal dependences that extend over an extended period of time, for the reasons that we will understand in Section \textbf{faltacitação}. Thus, in 1997 Hochreiter et al. proposed a novel RNN structure~\cite{Hoch97}, the Long Short-Term Network (LSTM), where a memory cell was introduced, and the input/output/read/write access are controlled by individual gates that are activated both by the incoming data samples, but also by the output from the previous time-step (it is an RNN after all). They are one of the state of the art methods in Deep Learning nowadays, as we can attest in Section~\ref{sec:sa_apps}.

\section{Motivation}\label{sec:intro_mot}
Hitherto, and to the best of my knowledge, all of the published applications that use LSTM are software based, but the parallel nature of the structure hollers for a dedicated hardware realization, that can dramatically increase its performance, something that, hitherto, has only recently been done once~\cite{Chang15} (see Section~\ref{sec:sa_hardware} for further details), and although it improves the processing time when compared to a naïve software solution, it still lacks the ability to perform on-chip learning, and the learning process still needs to be performed offline, in a normal CPU, when it also could be speeded up in a dedicated hardware structure.

All these techniques are generally implemented in mainstream processors, making use of general high-level or low-level languages, where all the real parallelism is limited to the number of simultaneous threads that we can run on each physical core, which up to the date generally have between 2-8 cores (mobile devices and general use personal computers),

In order to parallelize the computations to the fullest extent possible, a solution is to port it both to a Graphics Processing Unit (GPU) or even a Field-Programmable Gate Array (FPGA), but the portin process is not entierely automatic, and to have the least performance drop possible, it has to be explicitly programmed in CUDA/OpenCL for GPUs, and a Hardware Description Language for FPGAs, with an increasing level of complexity and low-level details. This way, it is necessary to provide frameworks that can allow an FPGA to quickly reconfigure itself to run these kind of networks on demand, for a particular task that requires them, achieving a supposedly lower computation time, and unburdening the CPU from running it, and saving performance room for running other tasks related, for instance, with the Operating System. Furthermore, when processing an incoming stream of highly dimensional data, or with high throughput, a CPU solution might not be scalable, and could benefit greatly from a dedicated hardware implementation.

\section{Objectives}\label{sec:intro_obj}
Taking into account the considerations done in~\ref{sec:intro_mot}, I propose to develop a hardware implementation of an LSTM Network, \emph{with on-chip learning}, improving the performance, capability and flexibility of the existing solution of~\cite{Chang15}. Moreover, I also propose to benchmark this solution and try to compare it with the software performance results of~\cite{Tavear13} and, as a secondary objective, try to use my developed structure to replicate the some of the applications portrayed in~\ref{sec:sa_apps}. Lastly, to make justice to the name of the thesis, I will work to make this structure reconfigurable on the go, minimizing the reconfiguration time.

\section{People Involved}\label{sec:intro_people}
Besides me, the candidate to the Master's Degree, there are two more people involved, namely

\begin{itemize}
    \item \textbf{Supervisor} -- The \href{https://sigarra.up.pt/feup/pt/func\_geral.formview?p\_codigo=210963}{Professor João Canas Ferreira}, auxiliary professor at the Faculty of Engineering of the University of Porto.
    \item \textbf{Second Supervisor} -- \href{http://www.cl.cam.ac.uk/~ijpdmt2/}{Ivo Timóteo}, a Computer Science PhD candidate at Cambridge University, UK, in the field of Artificial Inteligence.
\end{itemize}