### 2016-02-12 prepReport/earlyWork.tex
Applied all improvements suggested

\chapter{Early Work}\label{chap:earlyWork}
In order to better undesrtand LSTM,  gain some useful insight on their operation, and to have a software implementation, that is not part of any framework, with which I can validate in the future my FPGA implementation, I have developed a fully functional LSTM Layer in Python and Numpy, that still uses BPTT as a training algorithm (the governing equations are in appendix A of~\cite{Greff15}). The idea is to adapt it to use SPSA in the beginning of my Master's work. The code is present in

\lstinputlisting[language=Python, caption=Implementation of a LSTM Layer]{LSTMlayer.py}