### 2016-07-01 finalText/proposedSol.tex

\chapter{Proposed Architecture}\label{chap:propSol}

\section{Constituent Modules}

\subsection{Finite-precision representation system}
Before discussing any other details concerned with the actual hardware implementation, I will layout some of the design choices that were made regarding how the numbers involved in the calculations are represented. For that purpose, Section~\ref{sec:precbit} discusses the fundamentals of fixed-point representation systems, as well as the bitwidth and precision chosen for then number representation system of the network, and Section~\ref{sec:convrulesfp} states how the conversion between real and fixed-point numbers is performed. Finally, Section~\ref{sec:arithrulesfp} explains the special rules that fixed-point arithmetic imposes. The considerations made in this chapter can be found on~\cite{Yates13}, which is a good reference for fixed-point arithmetic theory.

\subsubsection{Precision bitwidth}\label{sec:precbit} 
Since we are dealing with real numbers, and I plan to make use of the DSP48 slices within the FPGA, I chose to use an 18-bit signed fixed-point system, with the sign information coded as 2's complement. Fixed-point systems are usually addressed in the $Qn.m$ form, where $n$ is the bitwidth of the of the integer part (excluding the sign bit) and $m$ is the bitwidth of the fractional part, and so the total bitwidth is $N=m+n+1$ to account for the sign bit. In this way, the value of the $i$-th position bit is $2^{i-m}$, and since this is a 2's complement system, the last bit is $-2^{n-1}$. In terms of range of representation, the maximum positive number that can be represented corresponds to all bits set to one, except for the last ($2^{n+m+1}-1$), but shifted by $m$ bits to the right to yield the correct real number (the decimal point is at the $m$-th bit),

\begin{equation}\label{eq:maxpos}
    \text{Max. Positive Number} = \frac{2^{N-1}-1}{2^m} = 2^{N-1-m} - \frac{1}{2^m}
\end{equation}
and the smallest negative number is simply the MSB set to one (and also shifted appropriately),

\begin{equation}\label{eq:minneg}
    \text{Small. Negative Number} = -\frac{2^{N-1}}{2^m} = -2^{N-1-m}
\end{equation}

Assuming that the smallest perturbation used in the training system will be $2^{-9}$, according to the previous Python experiments, our fractional part precision should be, at least, greater than this, so a sensible choice would be either $m=10$ or $m=11$. On the other hand, given that in the Python experiments I have attested that the weight values generally do not (and should not) exceed values around the first power of ten, there is no need for large values of $n$ (although it is advisable to be large enough to accommodate the intermediate calculations, avoiding overflow), and so we should choose to have as much precision as possible, so $m=11$. Since $18=n+m+1$, then $n = 17 - m = 6$ and the representation system to be used is $Q6.11$. According to Equations~\ref{eq:maxpos} and~\ref{eq:minneg}, the real values $x$ than can be represented with this choice of $n$ and $m$ are in the range

\begin{equation}\label{eq:rangeQ611}
    -64 \leq x \leq 63.99951172
\end{equation}
with a minimum resolution of $2^{-11} = 0.00048828125$.

\subsubsection{Conversion between real and fixed-point}\label{sec:convrulesfp}
In order to find the real number equivalent of a $Qn.m$ fixed-point system, and \textit{vice-versa}, we need to take into account that, according to Section~\ref{sec:precbit}, the $i$-th position bit is worth $2^{i-m}$, and therefore the decimal point in this fixed-point system is at position $i=m$, since $2^{m-m} = 1$. This way, the rules are as follows

\begin{itemize}
    \item \textbf{Positive Real to $Qn.m$} -- since the $1$ is at bit $m$, we simply multiply the real number by $2^m$, and discard the fractional part of the result
    \item \textbf{Negative Real to $Qn.m$} -- we disregard the sign in the real number, and perform the same operation as before, but we then convert the resulting binary number to two's complement, i.e. by performing bitwise negation, followed by summing 1.
    \item \textbf{Positive $Qn.m$ to real} -- multiply the fixed-point number by $2^{-m}$, to shift the decimal point back by $m$ positions. 
    \item \textbf{Negative $Qn.m$ to real} -- convert from two's complement by performing bitwise logic negation, followed by summing one; then scale the decimal point back by $m$ positions by multiplying by $2^{-m}$. 
\end{itemize}

\subsubsection{Fixed-point arithmetic rules}\label{sec:arithrulesfp}
The three main operations needed in my network design are \textbf{signed sums}, \textbf{signed multiplications} and \textbf{arithmetic shifts} (i.e. the ones that preserve the sign of the MSB) to implement divisions/multiplications by powers of two. In terms of signed sums, the rule is simple: both numbers need to be scaled to the same base, with their $m$'s being same, so that the decimal point is in the same place in both numbers ($n$, however, can be different, since that only means that one number is longer than the other, and the missing bits in the smaller one can be interpreted as zeros).

For signed multiplication, since both operands are in fixed-point $Qn.m$, and are thus scaled by $2^m$, we need to scale the result by multiplying it by $2^{-m}$ (or perform an arithmetic right shift of $m$ bits). This is because, if $a$ and $b$ the real numbers to be multiplied, we have that in fixed-point arithmetic, the result $c$ is

\begin{equation}\label{eq:multfp}
    (a\cdot2^m) \cdot (b\cdot2^m) = c \cdot 2^{2m}.
\end{equation}
Since in $Qn.m$, all numbers are represented as $r 2^{m}$ with respect to their real counterpart $r$, we need to scale back the decimal point in the result of Equation~\ref{eq:multfp}. We can see that it can be easily achieved by dividing by $2^m$, as stated in the previous paragraph.

\subsection{Non-linearity Calculator}\label{sec:nonlincalc}
In order to evaluate the non-linear activation functions $\sigma(x)$ and $\tanh(x)$, since there is no algorithm that can directly compute them, I had to find a suitable way to compute them accurately using a finite number of elementary operations, multiplications and additions, that can be performed efficiently by specially tailored blocks within the FPGA (DSP48 slices for multiplication, for instance). For that purpose, after using~\cite{Muller05} as a reference on elementary function approximation algorithms, I decided to use \textbf{Polynomial Approximations}, since evaluating a polynomial does not have high memory usage needs (as opposed to Table Methods, for instance) and, if the polynomial degree is sufficiently low, the number of multiplications needed is low enough to not pose a restriction both on resources (now DSP slices, and not memory) and in speed (number of number of clock cycles needed to output a result).

\subsubsection{Theoretical considerations on the approximation method}
The polynomial approximation methods aim to approximate some function $f(x)$ in an interval $\left[a, b\right]$ using a polynomial $p^*_n \in \mathcal{P}$ of degree $n$, in order to meet the optimization criteria chosen \textit{a priori}. This optimization criteria can be either the well-known \textbf{Least Squares Approximation} procedure, where we minimize the \emph{average quadratic error} $\left[f(x)-p^*(x)\right]^2$, or the \textbf{Least Maximum Approximation}, where we \emph{minimize} the \emph{maximum possible error}, also commonly called a \textit{minimax} approximation. Since we are operating in $Q6.11$ fixed-point arithmetic, we need to guarantee that the maximum approximation error does not exceed the precision limit of this representation, $2^{-11}$, and so a \textit{minimax} approach is desirable, since it guarantees that a given maximum error is not exceeded. Weierstrass's Theorem~\cite{Weierstrass1885}, from 1885, guarantees that there is always a polynomial that can approximate any continuous function f with error $\epsilon > 0$. Chebyshev also proved~\cite{Muller05} that, in a \textit{minimax} polynomial approximation of degree $n$, the minimum approximation error $\epsilon$ is achieved in at least $n+2$ points, and the sign of approximation error alternates from one interval to the other, and thus the error is not \emph{biased}. This leads to a linear system of $n+2$ equations whose $i$-th line is given by

\begin{equation}\label{eq:remezline}
    p(x_i) - f(x_i) = (-1)^{n+1} \epsilon \Leftrightarrow p_0 + p_1 x_i + p_2 x_i^2 + \cdots + p_n x_i^n - f(x_i) = (-1)^{n+1} \epsilon
\end{equation}

The optimal coefficients of this \textit{minimax} polynomial are found using the \textit{Remez Algorithm}, which provides an iterative approach to solve the linear system given by Equation~\ref{eq:remezline} by finding, in each iteration, the $n+2$ set of points $x_i$ of Chebyshev's Theorem that minimize the error function to $\epsilon$. The algorithm operations are as follows

\begin{enumerate}
    \item Initializing the set $x_i$ of points to $x_i = \frac{a+b}{2} + \frac{(b-a)}{2}\cos\left(\frac{i\pi}{n+1}\right), \, 0 \leq i \leq n+1$ 
    \item Solve the system in~\ref{eq:remezline}
    \item Given the polynomial coefficients yielded by step 2, compute the $y_i$ points that minimize $p(x)-f(x)$, and replace the $x_i$s of the next iteration by these $y_i$s. Go to step 2 until $\epsilon$ does not decrease.
\end{enumerate}

On one hand, the system of~\ref{eq:remezline} of Step 2 can be written in matricial notation as

\begin{equation}
\begin{bmatrix} 1 & x_0 & x_0^2 & \cdots & x_0^n & - 1 \\ 1 & x_1 & x_1^2 & \cdots & x_1^n & + 1 \\  &  &  & \vdots & & \\ 1 & x_{n+1} & x_{n+1}^2 & \cdots & x_{n+1}^n & (-1)^{n+1} \end{bmatrix}
\begin{bmatrix} p_0 \\ p_1 \\ p_2 \\ \vdots \\ p_n \\ \epsilon \end{bmatrix} = \begin{bmatrix} f(x_0) \\ f(x_1) \\ \vdots \\ f(x_{n+1}) \end{bmatrix}
\end{equation}
and therefore the solution vector $p = \left[p_0 \, p_1 \, p_2 \, \cdots \, p_n \, \epsilon \right]$ for Step 2 of Remez's Algorithm is simply given by $p = A^{-1}b$, where $A$ is the matrix on the left-hand side of the equation and $b$ is the vector on the right-hand side. The Python implementation of this algorithm is presented in Listing~\ref{lst:remezpy} of Appendix~\ref{ap1:remez-pyth}.

Instead of using a \emph{single} polynomial of higher degree ($n \geq 3$) for the whole domain, I chose to partition the domain of the activation functions in \textbf{6 intervals}, and approximate each of those intervals using polynomials of degree $n=2$. This proved to yield a lower overall approximation error, as expected (the interval on which to perform the approximation is smaller). Also, since both the $\sigma(x)$ and $\tanh(x)$ have horizontal asymptotes in $\{0,1\}$ and $\{-1,1\}$ respectively, the far-left and far-right intervals do not need a polynomial approximation, and can be assigned a constant value equal to the corresponding asymptote. The resulting \textit{minimax} approximation polynomials yielded by running the code in Listing~\ref{lst:remezpy} are

\begin{equation}\label{eq:coefs_sigm}
\hat{\sigma(x)} = \left\{ 
\begin{array}{lc} 
0 & x  \leq -6 \\
0.20323428 + 0.0717631x + 0.00642858x^2 & -6 \leq x \leq -3 \\
0.50195831 + 0.27269294x + 0.04059181x^2 & -3 \leq  x \leq 0 \\
0.49805785 + 0.27266221x - 0.04058115x^2 &  0 \leq  x \leq 3 \\
0.7967568 + 0.07175359x - 0.00642671x^2 & 3 \leq  x \leq 6 \\
1 & x > 6
\end{array}
\right.
\end{equation}
for the sigmoid function and

\begin{equation}\label{eq:coefs_tanh}
\hat{\tanh(x)} = \left\{ 
\begin{array}{lc} 
-1 & x  \leq -3 \\
-0.39814608 + 0.46527859x + 0.09007576x^2 & -3 \leq x \leq -1 \\
0.0031444 + 1.08381219x + 0.31592922x^2 & -1 \leq  x \leq 0 \\
-0.00349517 + 1.08538355x -0.31676793x^2 &  0 \leq  x \leq 1 \\
0.39878032 + 0.46509003x - 0.09013554x^2 & 1 \leq  x \leq 3 \\
1 & x > 3
\end{array}
\right.
\end{equation}
for the hyperbolic tangent function. These constants were converted to $Qn.m$ using the rules in~\ref{sec:convrulesfp}, and embedded in the HDL model.

\subsubsection{Hardware Implementation}
The hardware module that implements these non-linearities is, essentially, a 2nd degree polynomial calculator, where the coefficients of the polynomial are chosen accordingly with the value of the input $x$. This last functionality
is implemented using a simple multiplexer that loads the signals $p_0$, $p_1$ and $p_2$ with the coefficients of Equations~\ref{eq:coefs_sigm} and~\ref{eq:coefs_tanh} based on the value of the input operand $x$. As for the polynomial calculator is concerned, although we could use a full-pipelined evaluator, that would require two DSP slices -- to perform the two simultaneous multiplications -- but the DSP slices will be heavily used in the matrix-vector calculators, so it is advisable to save them for that purpose. A simpler approach is to note that, according to Horner's Rule, we get

\begin{equation}\label{eq:factorPol}
p(x) = p_0 + p_1x + p_2x^2 = p_0 + x(p_1 + xp_2)
\end{equation}
where we can note that this operation can be divided in a two-step procedure of a simultaneous multiplication of the operand by a constant, and a subsequent addition of another constant: first, by multiplying the operand by $p_2$ and summing $p_1$, and then by multiplying the operand by this last result and summing $p_0$. The block diagram of the hardware implementation of the non-linearity calculator is presented in Figure~\ref{fig:nonlin}. Also, in Figure~\ref{fig:nonlin-out}, the output of the Verilog module that implements this design is compared with the actual output (I used Python3's Numpy as reference) for both activation functions.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/non-lin.eps}
    \caption[Block Diagram of the Non-linearity Calculator]{Block Diagram of the Non-linearity Calculator using a single multiplication. The multiplexer state is controlled by the flip-flip and sum block, that change state every clock cycle. When reset is applied, the selector is set to zero.}
    \label{fig:nonlin}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}{0.85\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figures/nonlin-out.eps}
        \caption[Plot of the output of the Sigmoid Calculator HDL module]{Plot of the output of the Sigmoid Calculator HDL module}
        \label{fig:nonlin-out-sig}
    \end{subfigure}
    \begin{subfigure}{0.85\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figures/nonlin-out-tanh.eps}
        \caption[Plot of the output of the Hyperbolic Tangent HDL module]{Plot of the output of the Hyperbolic Tangent HDL module}
        \label{fig:nonlin-out-tanh}
    \end{subfigure}
    \caption[The output of the HDL implementation of the activation functions]{The output of the HDL implementation of the activation functions}
    \label{fig:nonlin-out}
\end{figure}

\subsection{Matrix-vector Dot Product Unit}\label{sec:dotprod_sec}
From the set of Equations~\ref{eq:equationsLSTM}, we see that the weight matrices $\mb{W}_*$ and $\mb{R}_*$ are multiplied by the input vector $\mb{x}$ and the layer output vector $\mb{y}$, respectively. This way, we need a HDL block that implements matrix-vector multiplication, and that block must be parameterized in order to accommodate different matrices/vectors of various sizes: note that $x$ has length $M$ (the number of inputs to the layer), and so $\mb{W}_*$ has size $N\times M$, while $y$ has length $N$ (the number of neurons in the layer), and so $\mb{R}_*$ now has size $N\times N$. Plus, if we need a layer with different parameters, either in terms of number of inputs or number of neurons, it advisable to only change the respective parameter at the synthesis stage instead of having to redesign the whole block for that particular size.

The matrix-vector dot product of a matrix $A$ of size $N \times M$ by a vector $x$ of size $M$, if performed in a linear non-parallel way, can be described in terms of Algorithm~\ref{matvec-alg}

\begin{algorithm}
\begin{algorithmic}
\For {$i$ = $1:N$}
    \For {$j$ = $1:M$}
    \State $y_i := y_i + A_{ij} \cdot x_j$
    \EndFor
\EndFor
\end{algorithmic}
\caption{Matrix-vector multiplication of a matrix}
\label{matvec-alg}
\end{algorithm}
This operation has a computational complexity of $O(n^2)$. It can be seen that each of the $i$-th component of the output vector $y$ can be calculated \textbf{in parallel}, each only requiring the corresponding $i$-th line from the matrix. If we follow this approach, matrix-vector multiplication can now be performed in \textbf{linear time}, which is one of the great advantages of custom-tailored hardware solutions.

Although this solution only requires one multiplication per row of the input matrix (i.e. $N$ multiplications), if the row size is large, we may run out of resources in the FPGA; therefore, some sort of \textit{resource multiplexing} strategy must be used to ensure the flexibility of the solution to accomodate networks of larger dimensions. The solution I have found for this issue was to \emph{share} the multiplication slice between rows of the matrix: in a direct implementation of Algorithm~\ref{matvec-alg}, each multiplication slice was responsible for producing the $i$-th element of the output vector $y$ (of size $N$), therefore the final result for the vector would be ready in $M$ clock cycles (i.e. the number of columns); now, if we define a parameter $K_G = \frac{\text{Number of rows}}{\text{Number of multipliers}}$ -- the number of rows that share the same multiplier -- the same multiplier is responsible for producing several $i$-th elements of the output vector, in consecutive time slots of $M$ clock cycles. Suppose that we have a $8\times2$ matrix, and that $K_G = 2$; in this way, we would have 4 multipliers, and the output vector elements $y_0$, $y_2$, $y_4$ and $y_6$ would be ready after $M=2$ clock cycles, and the remaining -- $y_1$, $y_3$, $y_5$ and $y_7$ --  are ready after another two clock cycles, that is $2M = 4$ clock cycles after the calculations began.

Figures-\ref{fig:mem-arrayprod} and~\ref{fig:array-prod} depict a diagram of the memory access for the Matrix, and the row multiplication and units within the module, respectively, where I have set $K_G = 4$, for the same matrix and vector sizes as before. Note that in this situation, we would only have 2 multipliers, and the module would be composed of two multiplication units, such as those in Figure~\ref{fig:array-prod}, that work in parallel: they address a particular column using the signal \verb+colSel+, which is used by the RAM module to output the corresponding column of the matrix (in regard to the input vector, obviously this signal selects only a single position), depicted in Figure~\ref{fig:mem-arrayprod}. The dark shaded part of the memory is used by the first multiplier, and the light shaded is used by the other, in parallel for a fixed \verb+rowMux+ -- this signal is produced by the control unit of the module, and essentially operates the left Mux and right Demux of Figure~\ref{fig:array-prod} that allows to choose the proper position of the weight column and to write to the correct output register, respectively. In this example, for \verb+rowMux+=0, the control unit increments \verb+colSel+ from 0 to $M$, and thus evaluating $y_0$ and $y_4$. After this, \verb+rowMux+ is incremented to 1, and the process repeats until \verb+rowMux+ reaches $K_G-1$. Therefore, we have the correct result vector in a time proportional to $K_G \cdot M$: since the memory only outputs the appropriate column in the \emph{next clock cycle}, and that this module is pipelined both at the input and at the output in order to increase the maximum clock frequency, we should add a 3 clock cycle overhead to the previous estimate.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/mem-array-prod.eps}
    \caption[A column of the matrix that serves as input to the module]{A column of the matrix that serves as input to the module. The dark shaded part is for the first multiplier, and the light shaded is for the other, in parallel. The rowMux signal addresses the position within each shaded area}.
    \label{fig:mem-arrayprod}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/array-prod.eps}
    \caption[The $i$-th row multiplication unit of the Module]{The $i$-th row multiplication unit of the Module, where rowMux and colSel are internal signals produced by the control unit of the module. The flip-flop accumulates the sum, and the output demux selects the appropriate memory position on where to store this value, within the slot attributed to this multiplication, from $i\cdot K_G$ to $i\cdot K_G + \left[K_G-1\right]$}
    \label{fig:array-prod}
\end{figure}

\subsection{Dual-port weight RAM}
Even though I have chosen to keep the bias weight vectors in normal registers and let the synthesis tool decide how to place them in the FPGA, for the weight matrices, the case is different. Since the network size can be large, it is good to make sure that they are placed in the RAMs available in the FPGA. In Section~\ref{sec:dotprod_sec}, we see that, because of the architecture of the matrix-vector multiplier, it is convenient to have a column by column access to the weight matrix, and therefore this block outputs a given column in the \verb+rowOut+ output port terminal, at the \emph{negative edge} of the clock, selected by the value present in the input \verb+addressOut+ at the immediately previous \emph{positive edge} of the clock. As far as \textbf{writing} to the memory is concerned, the process is identical but now the column of weights at the input \verb+rowIn+ is sampled at \emph{negative edge} of the clock, and is placed at the column specified by the input \verb+addressIn+ at the previous \emph{positive edge} of the clock.

Note that, for an $N \times M$ matrix, since the memory outputs and writes one column at a time, both the input and output port terminals will carry $N$ weights, and a total bitwidth of $\verb+BITWIDTH+ \cdot N$.

The Verilog coding followed the Verilog Coding Guidelines in Xilinx's \href{http://www.xilinx.com/support/documentation/sw_manuals/xilinx2015_2/ug901-vivado-synthesis.pdf}{UG901} that, in page 50, recommends that the \verb+RAM_STYLE+ parameter be set to \verb+block+. This can be done by adding the follwing compiler directive before the register definition, as follows

\begin{verbatim}
(* ram_style = "block" *) reg [PORT_BITWIDTH-1:0] RAM_matrix [NCOL-1:0];
\end{verbatim}
where we can see that each register contains the respective column of the matrix, with $\verb+PORT_BITWIDTH+ = \verb+BITWDITH+ \cdot M$ and $\verb+NCOL+ = M$.

After performing synthesis, I have noticed that using block RAM led to a slightly poorer performance, so I have settled with the use of \emph{Distributed RAM}, storing the weight matrices in LUTRAM.

\subsection{Gate Module}\label{sec:gatemod} 
The Gate modules are responsible for producing the internal signal vectors for $\mb{z}^{(t)}$, $\mb{i}^{(t)}$, $\mb{f}^{(t)}$ and $\mb{o}^{(t)}$. If we note that, according to~\cite{Greff15}, the removal of peepholes (the signals $\mb{p}_*$) does not compromise significantly the performance of the network~\ref{sec:struct_lstm}, we can ommit them in order to simplify the Gate Module and reduce the usage of DSP slices. This way, a Gate module needs to perform three tasks

\begin{enumerate}
    \item Multiply matrix $\mb{W}_*$ by the input vector $\mb{x}^{(t)}$
    \item Multiply matrix $\mb{R}_*$ by the previous layer output vector $\mb{y}^{(t-1)}$
    \item Sum the bias vector $\mb{b}_*$ to the remaining matrix-vector dot product results.
\end{enumerate}
Assuming that the network size $N$ is always larger than the input size $M$, if we use the matrix-vector dot product units of Section~\ref{sec:dotprod_sec}, the multiplication in task 1 takes approximately $K_G\cdot M$ cycles and the one in task 2 takes $K_G\cdot N$ cycles. This way, task 2 and task 1 can be performed in parallel, and we can use the extra time that task 2 takes, relative to task 1,  to perform task 3, and sum the bias vector to the output of task 1, whose result is ready by that time. The module is triggered by a \verb+beginCalc+ input signal that activates the internal state machine, and outputs a \verb+dataReady+ signal that informs the network that the calculations have been concluded, and the value at the output is the actual final result. 
After validating and simulating this block, and taking into account the internal state-machine and that the internal datapath is pipelined, the exact number of clock cycles this module takes to output a result is $6+K_G\cdot N$.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/gate.eps}
    \caption[Diagram of the hardware block that implements the Gate]{Diagram of the hardware block that implements the Gate}
    \label{fig:gate}
\end{figure}

\section{Hardware LSTM Network -- Forward Propagation}\label{sec:hLSTM-fp}
After detailing the design of each of the constituent modules of the LSTM Network, I will now show how they are interconnected
to implement a hardware version of Equations~\ref{eq:equationsLSTM}. This is performed in Section~\ref{sec:hLSTM-struct}, where the design decisions and compromises that were taken are detailed, along with some estimates on the resource usage and calculation time for each of the design iteration versions, the naïve solution (~\ref{sec:struct-fullpar}) and an optimized version of the former (~\ref{sec:struct-shared}) that exploits some of the hardware redundancies by multiplexing in time the usage of those redundant structures.

\subsection{Network Structure}\label{sec:hLSTM-struct}
Looking at equations~\ref{eq:equationsLSTM}, we note that the signals $\mb{z}^{(t)}$, $\mb{i}^{(t)}$, $\mb{f}^{(t)}$ and $\mb{o}^{(t)}$ do not depend on each other -- they operate only on the current input vector $\mb{x}^{(t)}$ and the previous layer output $\mb{y}^{(t-1)}$ -- we see that they can be calculated in parallel, meaning that we need four Gate Modules (see Section~\ref{sec:gatemod} working in parallel, each one with its respective two Block RAMs for $\mb{W}_*$ and $\mb{R}_*$, each followed by the respective activation function calculator (detailed in Section~\ref{sec:nonlincalc}). There are three elementwise multiplications, two for producing signal $\mb{c}^{(t)}$ (which can be done in parallel and then summed elementwise) and one for $\mb{y}^{(t)}$ (which can be done only after applying the activation function $\mb{c}^{(t)}$).

\subsubsection{Fully Parallel}\label{sec:struct-fullpar}
By implementing directly the ideas outlined previously, we get the network of Figure~\ref{fig:network}. The memory element of the network is the array of flip-flops that keep the value of the vector $\mb{c}^{(t)}$, which is activated everytime we have a new incoming signal in order to store the last value, which is now $\mb{c}^{(t)}$

(Talk about RAM usage)

In terms of DSP slice usage, this proposed design comprises 3 elementwise multiplications and 5 activation function Modules. Since the number of elementwise multiplications is equal to the network size $N$, we need $3N$ DSP slices for each elementwise block. As far as the activation function module is concerned, since we need to apply it to each one of the elements of the layer, $N$, and each module uses one multiplication, the ensemble of all 5 modules needs $5N$ DSP slices. Therefore, and noting that each Gate has $\frac{2N}{K_G}$ DSP slices, the total number of DSP slices used is

\begin{equation}\label{eq:numdsp_network}
    4\frac{2N}{K_G} + 5N + 3N = N \left( \frac{8}{K_G} + 8 \right).
\end{equation}
In terms of time performance, we can measure it by estimating the number of clock cycles needed to perform a complete forward propagation. Since the gate module outputs a result in $K_G \cdot N$ cycles, the activation function evaluators need always 5, and the elementwise multiplicators need only 2 cycles (one cycle for the pipeline, and another for the actual calculation), and noting that the two elementwise multiplicators that sum to produce $\mb{c}^{(t)}$ can work in parallel, the estimated number of clock cycles needed would be proportional to

\begin{equation}\label{eq:numcc_network}
    (N \cdot K_G + 6) + (5 + 2) + (5 + 2) = 20 + N\cdot K_G
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/network.eps}
    \caption[Hardware LSTM Network with full level of parallelism]{Hardware LSTM Network with full level of parallelism}
    \label{fig:network}
\end{figure}

\subsubsection{Shared Elementwise Multiplication Block}\label{sec:struct-shared}
In Figure~\ref{fig:network}, we see that, apart from the elementwise multiplier preceeding the flip-flops, all of them follow a similar structure: one of the operands is the output of a $\tanh(\mb{x})$ block and the other from a $\sigma(\mb{x})$. An improvement over the last network architecture is to instead of replicating these $\tanh$-$\sigma$-$(\cdot)$wise structures use a \emph{single one} and choose the operand accordingly to the state that the network is currently in. Besides, the right elementwise multiplier of Figure~\ref{fig:network} is not used as the same time as any other, so it is a perfect waste of resources. The issue about the elementwise multiplier that precedes the flip-flops can be solved by adding another multiplexer that chooses between the output of the $\tanh(\mb{x})$ module or the signal $\mb{c}^{(t-1)}$. These ideas resulted in the improved network design of Figure~\ref{fig:network-opt}.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/network-opt.eps}
    \caption[Optimised version of~\ref{fig:network} that shares the elementwise multiplication and the activation function blocks]{Optimised version of~\ref{fig:network} that shares the elementwise multiplication and the activation function blocks}
    \label{fig:network-opt}
\end{figure}
The two left multiplexers control the operands that are fed to the activation function modules, and the selecting signal is generated by the network's state machine, and its value is incremented after each complete usage of the $\tanh$-$\sigma$-$(\cdot)$wise  structure: this is where the time multiplexing of the structure takes place. Since in state $\verb+Sel+=1$ the left operand of the elementwise multiplicator (the one that preceded the flip-flops in the previous design) is the signal $\mb{c}^{(t-1)}$, I added another multiplexer before the elementwise multiplication that selects that signal in that particular case, and the output from the $\tanh(\mb{x})$ block, otherwise.

The flip-flops on the right hand side of Figure~\ref{fig:network-opt} are activated by signals generated within the network state machine that enable the appropriate flip-flop, placing the output from the elementwise multiplicator in the correct place. The first activated flip-flop is the middle one, which keeps the result from the operation of the $\mb{z}^{(t)}$ and $\mb{i}^{(t)}$ vectors, then, after a full operation of the elementwise multiplier, the bottom flip-flop to save the other portion of the sum that evaluates to the $\mb{c}^{(t)}$ signal. Lastly, the top flip-flop saves the network output $\mb{y}^{(t)}$, which in the next incoming sample becomes $\mb{y}^{(t-1)}$ and is used by the Gate modules in this next batch of calculations.

Now, since there is only a single elementwise multiplier and only two activation function calculators, the total requirement for DSP slices is simply
\begin{equation}\label{eq:numdsp_network-opt}
    4\frac{2N}{K_G} + 2N + N = N \left( \frac{8}{K_G} + 3 \right).
\end{equation}
where we see that we saved $5N$ multipliers, which for a large value of $N$ can have a decisive impact. In terms of speed performance, although the Gate calculation time remains the same, now the $\tanh$-$\sigma$-$(\cdot)$wise structure runs for 3 consecutive times, in a non parallel fashion. After adjustments to the state machine, and accounting for pipelining and synchronization within the datapath, the clock cycles needed after the gate calculations are 27, so the total clock cycles needed are

\begin{equation}\label{eq:numcc_network-opt}
    (N \cdot K_G + 6) + 27  = 33 + N\cdot K_G
\end{equation}
which is only 13 clock cycles more than the fully-parallel architecture. For instance, an $N=32$ neuron network would require 320 DSP slices, while this new architecture would only require 160, only at the expense of 13 more clock cycles.

\section{Hardware LSTM Network -- SPSA Training}
In this section, I propose a digital circuit that aims to train the Network whose details were explained in the previous Section by implementing the training algorithm of Section~\ref{sec:training_lstm}. The training consists in changing the weight values in the RAMs accordingly with the error of the current prediction against the correct outcome. This envolves \textbf{two} forward propagations: one, where we run the network with the weights unchanged, and another where we \emph{perturbate} the weights by a parameter $\pm \beta$ (the actual sign for each weight is generated at the beginning of the first forward propagation). After these two forward propagations, there is a small time frame where the weights are updated, row by row, before the next incoming sample. This is explained in Section~\ref{sec:train-circ}. The way how the numbers are randomly generated in hardware is presented in Section~\ref{sec:randnum-circ}.

\subsection{Pseudo-random Number Generator}\label{sec:randnum-circ}
While a True Random Number Generator (TRNG) generates data from a physical entropy source, a Pseudo-random Number Generator (PRNG) generates a sequence of numbers by applying a given
function to an initial seed. This means that, if one knows that seed and the function that generates the sequence, one can predict the previous and future numbers given the current state. While this
is not acceptable for high security cryptographic systems, it serves the purpose of providing a sequence of random sign vectors (from whom both the signal of the perturbation and the update will be choosen from).
The remaining issue is to find a PRNG with good statistical qualities, so instead of implementing a simple linear-feedback shift register (LFSR), I did some research on the topic, and eventually chose the solution
of~\cite{Tkacik2003}, based on a 43-bit LFSR and a 37-bit Cellular Automata, by selecting 32 bits from each, permuting the bits and performing a bitwise XOR -- the output has, also, 32 bits. The quality of the 
statistical properties are guaranteed, since it passed the DIEHARD test battery, as well as others, according to~\cite{Tkacik2003}.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/prng.eps}
    \caption[The 32 bit Pseudo-random Number Generator. Source~\cite{Tkacik2003}]{The 32 bit Pseudo-random Number Generator. Source~\cite{Tkacik2003}}
    \label{fig:prng}
\end{figure}
The generator polynomial of the LFSR is a maximal length polynomial, $x^{43}+x^{41}+x^{20}+x+1$, and thus has a period equal to $2^{43}-1$. The CA is also a register, but each bit is updated
according to a given \textbf{rule} based on the values of the neighboring bits. This particular CA follows Rule 150 for the bit in position $i=27$, given by $c_{t+1}(i) = c_{t}(i-1) + c_{t}(i) + c_{t}(i+1)$, and
the remaining bits are updated by Rule 90, given by $c_{t+1}(i) = c_{t}(i-1) + c_{t}(i+1)$. The bits positioned at the borders also follow rule 90, but treat as zeros the non-existing $i$ positions.

The output of my hardware implementation of this PRNG is presented in Figure~\ref{fig:prng_out}. It is a state-time diagram, where the contents of the registers at each time are represented by black (ones) and white (zeros), and the left column represents the output of the LFSR, the center column is the output of the Cellular Automata and the right column is their combined output. As one can see, the time dependence of the LFSR is very high, which is not desirable. The Cellular Automata, although uncorrelated, still retains some of the patterns that emerge from CAs. On the other hand, the combined output is the one that most resembles white noise, which is a desirable characteristic.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/prng_out.eps}
    \caption[The output of the implementation of the presented PRNG (right). The LFSR is the left column, and the CA is at the center]{The output of the implementation of the presented PRNG (right). The LFSR is the left column, and the CA is at the center. The zeros are represented by white, and the ones by black}
    \label{fig:prng_out}
\end{figure}

\subsection{Training Algorithm Circuit}\label{sec:train-circ}
Although the convergence results are not conclusive, here is presented the proposed digital circuit of the SPSA training for the network. The circuit of Figure~\ref{fig:train} is replicated for each of the eight weight RAMs, and the Sign RAM holds a single bit for each of the weight positions of the main weight RAM. This bit keeps the sign of the perturbation to be applied to that particular weight, both on the second forward propagation and the training step. During the first forward propagation, and since the weights are not perturbed in this run, we can use this time to propagate the sign RAM with new values for the next training stage. For this purpose, the PRNG is activated by the signal \verb+genRandNum+, generated by the state machine, and we write columns of bits to the sign RAM, re-using the addressing signal produced by the Gate, namely \verb+colAddressRead+ (\verb+weightUpdate+ is set to zero at this stage).
At the second forward propagation, \verb+pertWeights+ is set to one, and the Perturbation Mux selects the input where the weights are summed with the perturbation $\beta$, whose sign is selected by the sign RAM.

After this stage is completed, and before we start this new training cycle for the new incoming sample, we perform the actual training. For that, \verb+weightUpdate+ is set to one, and we enable writing to the Weight RAM. Also, the machine state is now responsible for addressing both RAMs, and we begin a simultaneous \emph{pipelined} write to the Weight RAM: this is done by delaying the the Write address in respect to the Read address. This way, at the first clock cyle while we read Column 0, in the next clock cycle we will read Column 1 \emph{and} write the sign-changed column 0 plus the update coming from the difference between the outputs in the two forward propagations. In order to simplify the update rule of Equation~\ref{eq:spsa_updateRule}, I have fixed $\alpha$ and $\beta$ as negative powers of two, and this replaces the multiplication and division by a simple arithmetic right shift -- this idea is also proposed by~\cite{Maeda05}.

The learning, although working in the Python model developed, is still not converging in the simulations of the Verilog code for this circuit. This might have to deal with an overflow problem at the weight update. This is a good question to be addressed in future work.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/train.eps}
    \caption[Proposed Hardware Implementation of SPSA Training]{Proposed Hardware Implementation of SPSA Training}
    \label{fig:train}
\end{figure}