### 2016-07-01 finalText/results.tex

\chapter{Results}\label{chap:results}
The optimized design of Section~\ref{sec:struct-shared} was synthesized for different network parameters, using Xilinx's Vivado 2015.4 Design Suite, and the synthesis results are reported in Section~\ref{sec:res-synth}, along with their discussion. The performance of the synthesized networks is very promising, showing an approximate $195\times$ speed-up over the LSTM Python model, which was running on a high-end desktop computer with an over clocked CPU. Although the FPGA aims for an Embedded System environment usage, the fairness of this comparison still holds, since a Python/Numpy implementation is slower than a lower-level language implementation (such as in C) and this level of performance would be similar to a C implementation in a low-power, embedded CPU platform. These results are reported in Section~\ref{sec:res-valid}.

\section{Synthesis}\label{sec:res-synth}
In order to assess the impact of the network size on the performance and resource usage of the proposed design, I have synthesized the Verilog description of the proposed network for sizes $N$ of 4, 8, 16 and 32, and for each of these the resource sharing parameter $K_G$ was also varied (as will be explained further ahead, the number of inputs is set to $M=2$). Since performance is the key factor for using an FPGA implementation, the synthesis strategy used was the \verb+Flow_PerfOptimized_high+, and the implementation strategy was \verb+Performance_ExplorePostRoutePhysOpt+ to perform a post-route optimization stage that proved useful in many cases to meet the clock specifications when the first Place\&Route pass failed to meet them. For a network size of 32 and $K_G=8$, the LUT usage exceeded the LUT resources available in the FPGA, so only lower values of $K_G$ were successfully synthesized.

\subsection{Maximum Frequency}\label{sec:res-synth-maxfreq}
Using the synthesis strategy outlined in the previous paragraph, the maximum clock frequency that does not cause timing violations of any sort is plotted in Figure~\ref{fig:maxfreq}. For $N=4$, since there are only 4 rows to be multiplied, the maximum value of $K_G$ is 4, and hence no synthesis was performed for $K_G = 8$; also, since $K_G = 2$ and $N=32$ would use $32(8/2+3) = 224$ DSP slices, that exceeds the 220 slices available in the Zynq 7020, so there is no synthesis data for that value.

First, for all network sizes, it is clear that with increasing $K_G$, the maximum clock speed decreases. This means that there is a critical path in the Matrix-Vector multiplication unit, whose multiplexer becomes increasingly complex for higher values of $K_G$. On the other hand, when $K_G$ is the same, smaller networks are faster than larger networks. The fastest design is an $N=4$ and $K_G = 2$ network, with a clock frequency of $140.154$ MHz, and the slowest one is an $N=32$ and $K_G=4$ network, clocked at $90.498$ MHz. The reference design used for validation in Section~\ref{sec:res-valid} is an $N=8$ and $K_G=2$ network is clocked at $132.89$ MHz, which yields a clock period of $7.525$ ns.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/maxfreq.eps}
    \caption[The maximum achievable clock frequency for several Network Sizes $N$ and resource sharing level $K_G$]{The maximum achievable clock frequency for several Network Sizes $N$ and resource sharing level $K_G$}
    \label{fig:maxfreq}
\end{figure}

\subsection{DSP Slices Usage}\label{sec:res-synth-dsp}
The estimates made in Equation~\ref{eq:numdsp_network-opt} were proven to be accurate, as Figure~\ref{fig:dspused} confirms. The reference network design uses 56 DSP slices, which corresponds to $25.45\%$ of the total number of DSP slices available.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/dspuse.eps}
    \caption[The number of DSP slices used for several Network Sizes $N$ and resource sharing level $K_G$]{The number of DSP slices used for several Network Sizes $N$ and resource sharing level $K_G$}
    \label{fig:dspused}
\end{figure}

\subsection{Power Usage}\label{sec:res-synth-power}
Another important metric of the performance of a design is its \textbf{power consumption}. All designs yielded a constant baseline value for the \emph{static} power consumption of around 120 mW, and the power consumption reported in Figure~\ref{fig:power} refers to the \emph{total} power consumptions, i.e. both the baseline static power and the dynamic power consumption. It is clear that the smaller the network is, the less power is consumed, as one would expect. Furthermore, an increasing level of resource sharing yields a substantially lower power consumption figure: this is because we use less DSP slices when we increase $K_G$. Of course, even though the power consumption is lower in that case, that comes at the expense of a lower clock frequency and more clock cycles elapsed per forward propagation.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/power.eps}
    \caption[Power consumption estimates for several Network Sizes $N$ and resource sharing level $K_G$]{Power Consumption estimates for several Network Sizes $N$ and resource sharing level $K_G$}
    \label{fig:power}
\end{figure}

\subsection{Other Resources Usage}\label{sec:res-synth-otheres}
Besides DSP slices, it is also worth mentioning the usage of LUT, LUTRAMs and Flip-Flops. In Table~\ref{tab:lut}, the usage of LUTs is reported, where we can see that although there is not a clear trend on how the LUT usage varies with increasing $K_G$, it is clear, and expectable, that the LUT usage increases with the size of the network by an approximate factor of 2, from $N=2$ to $N=16$. As for $N=32$, the usage does not follow this apparent trend, and rises sharply to 91\%. For $K_G = 8$ the usage even surpasses the maximum amount of LUTs available.
\begin{table}
    \centering
  \begin{tabular}{ | l | c | c | c |}
    \hline
    & $K_G=2$  & $K_G=4$ & $K_G=8$ \\ 
    \hline
    $N=4$ & 6.87\% & 6.04\% & N.A. \\    
    \hline
    $N=8$ & 14.64\% & 13.03\% & 14.11\% \\    
    \hline
    $N=16$ & 28.97\% & 27.72\% & 29.85\% \\    
    \hline
    $N=32$ & N.A. & 91.09\% & N.D. \\    
\hline
  \end{tabular}
    \caption{LUT usage for different $N$ and $K_G$}
    \label{tab:lut}
\end{table}

As for LUT, the FF usage also scales accordingly to a $2\times$ factor. Furthermore the LUTRAM usage scales well and does not pose a limitation on the network size. The usage results are reported in Table~\ref{tab:ff}.

\begin{table}
    \centering
  \begin{tabular}{ | l | c | c |}
    \hline
         & LUTRAM  & FF  \\ 
    \hline
    $N=4$ & 0.18\%  & 3.39\% \\    
    \hline
    $N=8$ & 4.41\%  & 6.7\% \\    
    \hline
    $N=16$ & 8.83\%  & 13.36\% \\    
    \hline
    $N=32$ & 17.66\% &  26.45\% \\    
\hline
  \end{tabular}
    \caption{Flip-flop usage for different $N$ and $K_G$}
    \label{tab:ff}
\end{table}

\section{Validation and Comparison}\label{sec:res-valid}
Over the course of this Section, the functionality of the network is verified against the Python reference module of~\ref{sec:res-pyth} that I have developed as a reference, both for the forward propagation of the network, as well as for the training algorithm. The methodology of this procedure is outlined in Section~\ref{sec:res-meth}. On Section~\ref{sec:res-valid-fpps}, I will assess how the performance results of Section~\ref{sec:res-synth-maxfreq} translate into how many classifications per second this design achieves, and how that figure compares with the capabilities of the Python module.

\subsection{Reference Python Module}\label{sec:res-pyth}
Before performing the Verilog description of the network, I have tested the ideas outlined in Section~\ref{chap:theorBack} first by building a software version of an LSTM Network. Since I wanted to quickly test the ideas without much effort, I decided to use Python and Numpy rather than MATLAB, since the former have higher performance at the same level of code complexity. The code for this Python class is listed in Appendix~\ref{ap1:lstm-pyth}. It supports both the forward propagation of an already trained network, but also provides functions to train a model using the SPSA method outlined in Section~\ref{sec:training_lstm} and also Backpropagation Through-Time as presented in~\cite{Greff15}. As the reader can attest by running the code himself, the SPSA proves to be a valid approach for training these networks, achieving convergence for the Machine Learning task that will be presented in the next paragraph, although it converges slower than Backpropagation Through Time.

The learning problem presented to both the software and hardware network is the sum of two binary numbers of 8 bits. The bits of each number at position $i$ are input to the network as a vector, and the network outputs its prediction of the correct value of the $i$-th bit of the \emph{result} number. After the whole number number is processed, the memory cells of the LSTM network are reset and a new addition task can be presented to the network. Even though this seems a rather simple problem, it accounts for all the essential issues at which this network excels: first, this is a \emph{classification} problem in which the Machine Learning algorithm needs to output a prediction based on the input feature vector, and that prediction has to take into account the \emph{history} of predictions so far, because the current bit is the sum of not only the bits of the two operands, but also the \emph{carry} generated at the last few positions -- this is where the memory cells of this special Neural Network come into effect.

\subsection{Methodology}\label{sec:res-meth}
The Python testbench that both trains and verifies the predictions of the software network of Appendix~\ref{ap1:lstm-pyth} is listed in Appendix~\ref{ap1:testbench-pyth}. This script, after achieving convergence, stores the trained network weights in \href{https://docs.python.org/3/library/pickle.html}{Pickle} format. These weights are, then, converted to $Q6.11$ fixed-point format and directly loaded into the Verilog model by the Verilog testbench. The Verilog testbench then applies in order the bits of randomly generated numbers, captures the output bit and checks whether it is correct or not, and counts the number of incorrect bits.

This network has, then, two inputs ($M=2$) and eight neurons ($N=8$). Furthermore, the output of the eight LSTM neurons is reduced to a single output by means of a simple neuron, as the one described in Figure~\ref{fig:modelNeuron_a}. This introduces a further 5 clock cycle overhead that is not accounted in Section~\ref{sec:res-valid-fpps}, since it is not actually part of the LSTM Network and, although it was placed in series in the datapath, after the LSTM network, it can run in parallel with the LSTM network, introducing no overhead whatsoever.

A Python script was written (see Appendix~\ref{ap1:runtb-lstm}) to automatically compile the Verilog code, generate the input and output data, and call Questasim to run the simulation of the HDL code. Figure~\ref{fig:res-meth-wave} shows the waveform viewer window of a run of the Verilog testbench, and in Appendix~\ref{ap1:transcript-cmd} shows the transcript of a successful run of this testbench, proving that the learning ability of the Hardware network still retains the predicting capability of the software network.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/wave.png}
    \caption[Screenshot of the simulation run of a complete forward propagation for $N=8$ and $K_G=2$]{Screenshot of the simulation run of a complete forward propagation for $N=8$ and $K_G=2$. The period between the two time cursors is the same as the one reported in Table~\ref{tab:process-time}}
    \label{fig:res-meth-wave}
\end{figure}
\subsection{Performance}\label{sec:res-valid-fpps}
To evaluate the performance of the network in terms of how fast it can produce data, I have defined a metric based on how many predictions it can produce per second (i.e. produce a new result bit in the output sequence), in \textbf{millions}. The prediction time is the time elapsed since a new input vector is applied \emph{to} the moment the LSTM Network produces an output vector (as discussed previously, the five clock cycles of the simple neuron at the output are \textbf{not} considered). To do this, we multiply the number of clock cycles yielded by Equation~\ref{eq:numcc_network-opt} by the equivalent clock period from the synthesis clock report of Figure~\ref{fig:maxfreq}. This result is epitomized in Table~\ref{tab:process-time}, where the calculation time of the Python module forward propagation function is also included. This time was measured using the \href{https://docs.python.org/3.5/library/timeit.html}{\textbf{timeit}} module, that allows the evaluation of the execution time of small pieces of code as well as complete functions with arguments. The Python code was run on a Linux System, powered by an i7-3770k Intel Processor, running at 4.2GHz.

\begin{table}
    \centering
    \begin{tabular}{ | l | c | c | c | c | c | }
    \hline
    & $K_G=8$  & $K_G=4$ & $K_G=2$ & Python & Speed-up \\ 
    \hline
    $N=4$ & N.A.  & 360.15 ns  & 292.535 ns & 65 $\mu$s & $\times180$ \\   
    \hline 
    $N=8$ & 960.3 ns  & 494 ns  &  368.725 ns & 72 $\mu$s & $\times194$ \\    
    \hline
    $N=16$ & 1.715 $\mu$s  & 921.5 ns  & 518.05 ns & 96 $\mu$s & $\times185$ \\
    \hline
    $N=32$ & N.D. & 1.78 $\mu$s  & N.A. & 185 $\mu$s  &  $\times104$ \\    
\hline
  \end{tabular}
    \caption{Total processing time for a single forward propagation, for hardware networks of various $K_G$ and the Python software network of equivalent size. The speed-up figure is calculated between the Python processing time versus the best hardware network time of the same size}
    \label{tab:process-time}
\end{table}
The performance increase is impressive, even for the slowest of the designs (the $N=32$ network). The hardware network is, at best, $\times200$ faster than the software counterpart, and at worst $\times100$ faster. Also, it is noticeable that increasing the level of resource sharing increases the computation time, since the level of parallelism is lower.

 Since the previous values are the time needed for a \emph{single} forward propagation, to know how many forward propagations we can perform \emph{per second}, we only need to invert the previous values. These values are presented in Figure~\ref{fig:Mclass-psec}. While the $N=8$ and $K_G=2$ network is able to perform around 3.2 million predictions per second, the Python model can only output around 14 thousand predictions, which is a very significant result that proves the relevance of this implementation.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/Mclass-psec.eps}
    \caption[Millions of classifications per second of each design according to the network size $N$]{Millions of classifications per second of each design according to the network size $N$. The comparison is between the software Python model and 3 networks of different levels of resource sharing $K_G$.}
    \label{fig:Mclass-psec}
\end{figure}