### 2016-07-25 finalText/conclusion.tex

\chapter{Conclusion}\label{chap:conclusion}
The LSTM Hardware architecture presented surpassed the performance of the custom-built software implementation by $251\times$, at best,
and also the only current hardware implementation by $14\times$, and solely making use of internal FPGA
resources, achieving a higher level of parallelism. The higher levels of parallelism of this work are achieved at the cost of increasing
design complexity, which undermines its scalability for higher sized networks, unlike the implementation of Chang et al.~\cite{Chang15}.
On the other hand, the HDL description of this work is parameterized, and is thus very flexible for networks of any size, not requiring
a redesign of the system every time a different sized network is required. Furthermore, making use of internal memory makes it suitable
for including an on-chip learning system that can perform training on the network weights.

Given these results, this architecture places itself as the current state of the art system in LSTM Neural Networks
implementations in hardware, providing the most efficient implementation to date.

A possible follow-up of the work of this thesis, is the integration of the training circuit in Section~\ref{sec:train-circ}, providing
an LSTM Network with on-chip training capabilities.