### 2016-07-25 finalText/stateArt.tex

\chapter{State of the Art}\label{chap:stateArt}

Over the course of this chapter, I am going to present an overview of the most recent developments related to the work of this thesis, both in terms of existing dedicated hardware implementations~\ref{sec:sa_hardware}, the most relevant work in adapting suitable training algorithms to hardware~\ref{sec:sa_training} and also some of the most relevant applications of LSTM~\ref{sec:sa_apps}, which are not FPGA-based, but demonstrate how LSTM are useful by themselves, and how well it competes with other Machine Learning algorithms in terms of long time-series dependences in data.

\section{LSTM Applications (non-FPGA related)}\label{sec:sa_apps}
LSTM Networks are nowadays one of the state of the art algorithms in deep-learning, and their performance is superior to that of other kinds of RNNs and Hidden Markov Models, both of which are generally used to address the same set of problems where LSTM are employed, namely predicting and producing classification decisions from time-series data. A very comprehensive description of applications can be found in one of the initial authors webpage dedicated to the subject~\footnote{\href{http://people.idsia.ch/~juergen/rnn.html}{http://people.idsia.ch/~juergen/rnn.html}}. I will now enumerate some of the leading edge applications of LSTM.

\begin{itemize}
    \item \textbf{Handwriting Recognition} -- an LSTM-based network~\cite{Bertolami09}, submitted by a team of the Technische Universität München, won the 2009 ICDAR Handwriting Recognition Contest, achieving a recognition rate of up to 91\%~\cite{ICDAR09}. LSTMs have also proven to surpass HMM-based models in terms of optical character recognition of printed text, as~\cite{Breuel13} suggests.

    \item \textbf{Speech Recognition} -- an architecture\cite{Graves13} proposed by Graves et al. in 2013 achieved an astonishing 17.7\% of accuracy on the TIMT Phoneme Recognition Benchmark, which up to the date is a new record. Furthermore, it has also been used for large scale acoustic modelling of speech~\cite{Sak14}.

    \item \textbf{Handwriting Synthesis} -- a comprehensive study by Graves shows, among other sequence generation tasks such as Text Prediction, that use of LSTM to produce synthetic handwriting, that looks incredibly similar to human-produced handwriting~\cite{Graves13_2}.

    \item \textbf{Translation} -- LSTM was used by Sutskever et al. (Google) to perform sequence-to-sequence translation on the WMT'14 dataset, translating English to French with a close to state of the art score~\cite{Sustkever14}.

    \item \textbf{Biomedical Applications} -- this network architecture was used in a protein homology detection scheme~\cite{Hochreiter07} using the SCOP 1.53 benchmark dataset, displaying a good performance when compared to other methods. Similarly, a recent article from 2015 by Sønderby et al. suggested the use of standalone LSTM and also Convolutional LSTM to perform sub cellular localization of proteins, given solely their protein sequence, with an astounding accuracy of 0.902~\cite{Sonderby15}

    \item \textbf{Music Analysis and Composition} -- Lehner et al. proposed a low-latecy solution based on LSTM, suitable for real-time detection of a singing voice in music recordings~\cite{Lehner15} whose performance surpassed other baseline methods, with lower latency.
        In terms of music transcription from audio data, there is a study that proposes the use of LSTM cells to perform a transcription of piano recordings to musical notes~\cite{Bock12}, in order to automate music transcription. The model was trained with recordings of both acoustic pianos and synthesized pianos, and the labelling was performed using an associated MIDI file for each piece that was used in the training, showing promising results. \cite{Coca13} suggests the use of LSTM to perform autonomous computer music composition, and Eck and Schmidhüber proposed LSTM to perform Blues improvisation in~\cite{Eck02}.

    \item \textbf{Video and Image Analysis} -- Vinyals et al., at Google, proposed an LSTM network for \textbf{image captioning}, preceded by a Convolutional Neural Network (CNN) to apply preprocessing to the images~\cite{Vinyals14}. The LSTM works as \emph{sentence generator}, that captions the images with state of the art performance.
        Besides image captioning, video captioning is also an interesting topic. Venugopalan et al. recently proposed a CNN + LSTM architecture to translate video sequences to natural language~\cite{Donahue14}, using the Microsoft Research Video Description Corpus as a dataset.There are other similar studies that combine image and video captioning~\cite{Donahue14_2}.
\end{itemize}

\section{Hardware Implementations of LSTM}\label{sec:sa_hardware}
Hitherto, there is but one actual implementation of an LSTM network in hardware, published recently
(Januray 2016) by Chang et al.~\cite{Chang15} in the international Journal of Advanced Research in
Electrical, Electronics and Instrumentation Engineering. It consists on the proposal of an LSTM network with $N=128$ for dedicated hardware,
targeting a Xilinx\textregistered~ Zedboard implementation. It uses a character-level language model from Andrej
Karpathy~\footnote{\href{https://github.com/karpathy/char-rnn}{https://github.com/karpathy/char-rnn}}, written in
Lua using the Torch7 framework~\footnote{\href{http://torch.ch/}{http://torch.ch/}}.

Although the article is not clear on whether there is active learning by the ARM CPU -- the authors refer that the CPU
loads the weights before operation, and that it changes them during operation, although how and why that change is done
is not even clearly explained, neither mathematically nor conceptually --, \textbf{there is no on-chip learning module}
in the FPGA according to the description provided. For that reason, the system stores the weights in external memory,
and accesses them using Direct Memory Access (DMA). Furthermore, the level of parallelism is low, since for an
$N=128$ network, the design uses only $22.73\%$ of the DSP48's available in the board. This is because the matrix multiplications
are serialized, and only one DSP is used for each matrix-vector product (aside from other uses).

The implementation is not flexible, since there is nothing that suggests a parameterization of the HDL description.
Hence, it is only possible to synthesize networks of size $N=128$, meaning that there needs to be a redesign of the network
for other network sizes. On the other hand, this is a valid approach for special applications that require networks of
this size, and where we only need a co-processor for LSTM networks, and not a dedicated standalone system.

\section{Training Algorithms and Implementation}\label{sec:sa_training}
As stated in~\ref{sec:sa_hardware}, the work of~\cite{Chang15} does not feature on-chip learning at FPGA level, although there
are a handful proposed solutions for it in recent literature. I will particularly look into the ones that use SPSA (see
Section~\ref{sec:training_lstm}), since that is the training algorithm of choice for my proposed solution, and also to the
ones that particularly apply SPSA to the training of Neural Networks~\cite{Maeda05}.

The SPSA algorithm was initially published by Spall in~\cite{Spall98}, and its theoretical details, are outlined in
Section~\ref{sec:training_lstm}. As to its applications to the training of general Neural Networks, the earliest examples
come from 1995 and 1996 in~\cite{Maeda95, Cauwen96} where SPSA is used to train a VLSI \emph{Analog} Neural Networks, a time
where the memory resources of digital circuitry were limited, and so most of these structures were analog-based. Its adequacy
was also established for \textbf{control problems}, such as those proposed in~\cite{Figueiredo97}, where a Time Delay Neural
Network is used to control an unknown plant in a linear feedback system.

In 2005, Maeda and Wakamura published a proposed SPSA hardware implementation~\cite{Maeda05} to train an Hopfield Neural Network
in an FPGA (and thus a digital system), achieving promising results in an Altera FPGA. The article carefully delineates the approach
taken, and also the hardware architecture designed.

Furthermore, a 2013 article by Tavear et al.~\cite{Tavear13} proposes, for the first time, using SPSA to train LSTM Neurons, although
the article focuses on proving the suitability of SPSA to LSTM, and no actual hardware implementation is done or proposed. The authors
simply demonstrate the suitability using conceptual arguments and by building a software model of an SPSA-trained LSTM network, and by
comparing both the performance and computing speed of their model with the results achieved by Hochreiter et al. in~\cite{Hochreiter07}.
Since the forward phase in both regular LSTM and SPSA-trained LSTM is the same, the computation time suffers no performance penalty whatsoever
and the learning ability is preserved to a high degree, showing that SPSA is a valid alternative do BPTT and other similar and more common training schemes.

\section{Final Overview}\label{sec:overview}
As we could attest from this small literature survey, although there is already an hardware implementation of LSTM, there is still a good deal
of room for improvement by adding on chip-learning to the system, and also to improve its performance and flexibility.
Furthermore,~\cite{Tavear13} shows that, at least for that particular case, the LSTM network doesn't suffer a great performance impact from
using SPSA training, as opposed to the more common BPTT, and~\cite{Maeda05} showed that an
SPSA hardware implementation is feasible.